Conformational plasticity of phospholipid headgroups in simulations and experiments

### 2021-05-02 manuscriptPGPE.tex

\author{Antonio Pe{\'o}n}
\affiliation{Spain}

@@

MD simulations could be used to interpret the lipid conformational ensembles from experiments due to their direct connection to the measured C-H bond order parameters. To find the best models for this purpose, we have evaluated all the available MD simulation force fields of PC~\cite{botan15}, PS~\cite{antila19}, PE (Fig.~\ref{HGorderParametersPE}), and PG (Fig.~\ref{HGorderParametersPOPG}) lipids against NMR data in the NMRlipids project. The comparison of hundreds of simulation trajectories simulated with approximately 20 different force fields reveal that none of the current parameters can correctly capture the lipid conformational ensemble and reproduce the experimental C-H bond order parameters of all segments. However, the best parameters (CHARMM36) correctly capture the essential differences between PC, PE, PG and PS headgroup order parameters (Fig.~\ref{structures} B).

@@

Interpretation of experimental order parameters using MD simulations
in previous sections suggest that PC, PE, PG and PS lipid headgroups are very flexible,
allowing them to bind proteins in various different conformations.
To test this prediction, we analyzed the conformations of all protein-bound lipids 
from structures deposited in the PDB \cite{berman00}.
We found 311 PC, 394 PE, 154 PG, and 35 PS lipid conformations that were bound to different types of proteins and were determined as a part of protein structure using crystallography or cryo-EM. The full list of structures with bound lipids is available from {\it Data/PDBdistributions/pdbtable.pdf} in Ref.~\cite{NMRlipidsIVbgit}.

>>>>

\author{Antonio Pe{\'o}n}
\affiliation{CIQUP, Centro de Investigação em Qu{\'i}mica, Departamento de Qu{\'i}mica e Bioqu{\'i}mica, Faculdade de Ci{\^e}ncias, Universidade do Porto, Porto, Portugal}

@@

MD simulations could be used to interpret the lipid conformational ensembles from experiments due to their direct connection to the measured C-H bond order parameters. To find the best models for this purpose, we took advantage of the power of the NMRlipids project which has gathered a massive number of trajectories using open collaboration. We have thereby evaluated all the available MD simulation force fields of PC~\cite{botan15}, PS~\cite{antila19}, PE (Fig.~\ref{HGorderParametersPE}), and PG (Fig.~\ref{HGorderParametersPOPG}) lipids against NMR data in the NMRlipids project. The comparison of hundreds of simulation trajectories simulated with approximately 20 different force fields reveal that none of the current parameters can correctly capture the lipid conformational ensemble and reproduce the experimental C-H bond order parameters of all segments. However, the best parameters (CHARMM36) correctly capture the essential differences between PC, PE, PG and PS headgroup order parameters (Fig.~\ref{structures} B).

@@

Interpretation of experimental order parameters using MD simulations
in previous sections suggest that PC, PE, PG and PS lipid headgroups are very flexible,
allowing them to bind proteins in various different conformations.
To test this prediction, we analyzed the conformations of all protein-bound lipids 
from structures deposited in the PDB \cite{berman00}.
We found 311 PC, 394 PE, 154 PG, and 35 PS lipid conformations that were bound to different types of proteins 
(including both integral and peripheral membrane proteins) and were determined as a part of protein structure using crystallography or cryo-EM. The full list of structures with bound lipids is available from {\it Data/PDBdistributions/pdbtable.pdf} in Ref.~\cite{NMRlipidsIVbgit}.
